---
title: "Peer Group Analysis"
author: "Sarah Blackstone & Stephanie Stevens"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

##install.packages("tidyverse")
##install.packages("readxl")


library(tidyverse)
library(readxl)
```

# Prepare the Data

```{r}
## read in excel file with output from SQL Server
dat <- read_excel("N:/OP/OIR/OIR-Common/ADS Projects/IPEDS Peer Comparisons/Task Force Demo Apr 2024/knn_data.xlsx") #file path


#pull Institution names into an object (needed later)
names <- dat[,1] 

#remove names of institutions to exclude from analysis
dat <- dat %>% select(-instnm) #update with institution variable name

#scale data
dat <- as.data.frame(scale(dat))

#add already determined weight values by variable

 weighted_data <- dat %>%
 		mutate(rd_expend = rd_expend * 13.91,
 		percent_ug = percent_ug * 13.77,
 		total_enrollment = total_enrollment * 12.3, 
 		inst_faculty_fte = inst_faculty_fte * 10.94,
 		grad_rate = grad_rate * 9.15,
   	public_funding_per_fte = public_funding_per_fte * 11.62,
    percent_ft_inst_fac = percent_ft_inst_fac * 10.87,
    percent_ft_ug = percent_ft_ug * 9,
    percent_ug_pell = percent_ug_pell * 7.45,
 			)
```

# Distance Matrix
We are going to create a distance matrix showing the cumulative distance between each school. We can extract the first column, which will correspond to the first school listed in the original dataframe (JMU).

```{r}

#create a distance matrix of all variables
jmu_dist <- as.matrix(dist(weighted_data[, c(1:9)])) 

#extract the first column (which is JMU)
jmu_dist <- as.data.frame(jmu_dist[1, ]) 

#create a variable reflecting row index
jmu_dist <- jmu_dist %>% mutate(index = row_number()) 

#add a row number to the names file; we will use this to merge the institution names with the index and distance data
names <- names %>% mutate(rn = row_number()) 

#join the institution names with the distance matrix
peer_group <- jmu_dist %>% left_join(names, by = c("index" = "rn")) 

#order by overall distance to see the "nearest neighbors"
peer_group <- peer_group %>% arrange(`jmu_dist[1, ]`) 

#export
write_csv(peer_group, "Institutions Ranked for Peer Group.csv")
```

# Data for Visualization
We need to select the top n "neighbors" to create distances summaries for each variable to display in the radar plots.
```{r}


#select the 20 nearest neighbors plus JMU, which is row 1
top_withJMU <- peer_group[c(1:21), ] 

#add row index to the original weighted data frame
weighted_data <-weighted_data %>% mutate(rn = row_number()) 

#select the data for all variables from the weighted data frame for the observations in the top20
data_for_radar <- weighted_data %>% filter(rn %in% top_withJMU$index) 

#join this data frame with top 20 dataframe containing overall distance, index and institution name.
data_for_radar <- data_for_radar %>% left_join(top_withJMU, by = c("rn" = "index")) 

#select the institution name- this was we have the names stored in an object in the order they appear in the data we are using for variable group distances
top20names_withJMU <- data_for_radar$instnm  


```

## Distances for Individual Variables (for Radar Plot)

Apply the following function to each variable in the data_for_radar df, updating the column indicies and the variable name each time
```{r}
fun <- function(df, cols) {
  groupname <- as.data.frame(as.matrix(dist(df[, cols])))
  groupname <-  groupname[, 1]
  as.data.frame(groupname)
}

##Examples from JMU peer group analysis variables - change variable name and group name to fit your needs
rd_expend <- fun(data_for_radar, c(1:1))
rd_expend <- rd_expend %>% rename(`R&D Expenditures` = groupname)


percent_ug <- fun(data_for_radar, c(2:2))
percent_ug <- percent_ug %>% rename(`Percent of Undergrad` = groupname)

total_enrollment <- fun(data_for_radar, c(3:3))
total_enrollment <- total_enrollment %>% rename(`Total Enrollment` = groupname)

inst_faculty_fte <- fun(data_for_radar, c(4:4))
inst_faculty_fte <- inst_faculty_fte %>% rename(`Instructional Faculty FTE` = groupname)

grad_rate <- fun(data_for_radar, c(5:5))
grad_rate <- grad_rate %>% rename(`Graduation Rate` = groupname)

public_funding_per_fte <- fun(data_for_radar, c(6:6))
public_funding_per_fte <- public_funding_per_fte %>% rename(`Public Funding per FTE` = groupname)

percent_ft_inst_fac <- fun(data_for_radar, c(7:7))
percent_ft_inst_fac <- percent_ft_inst_fac %>% rename(`Percent of FT Inst Faculty` = groupname)

percent_ft_ug <- fun(data_for_radar, c(8:8))
percent_ft_ug <- percent_ft_ug %>% rename(`Percent of FT Undergrad` = groupname)

percent_ug_pell <- fun(data_for_radar, c(9:9))
percent_ug_pell <- percent_ug_pell %>% rename(`Percent of Pell Students` = groupname)
```


Bind the data frames together with the list of the top 20 names.

```{r}
final_data <- cbind(rd_expend, percent_ug, total_enrollment, inst_faculty_fte,grad_rate,public_funding_per_fte, percent_ft_inst_fac, percent_ft_ug,percent_ug_pell,  top20names_withJMU)

final_data <- final_data[c(2:21), ] #remove JMU from the list
```

# Reshape and export data
```{r}

## first variable : last variable
final_data <- final_data %>% 
  pivot_longer(`R&D Expenditures`:`Percent of Pell Students`, names_to = "Variable", values_to = "Distance")

write_csv(final_data, "Data Radar Plots TEST.csv")


```




